---
title: "Final_QDA"
author: "Yacheng Xiao"
date: "2023-03-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Quadratic Discriminant Analysis - Standardized**

```{r}
# Read in the datasets
train_qda <- read.csv("train_stand.csv")
train_qda <- subset(train_qda, select = -c(1))
test_qda <- read.csv("test_stand.csv")
test_qda <- subset(test_qda, select = -c(1))
```

```{r}
# Train QDA
library("MASS")
true_val = test_qda$classes
qda_stand = qda(classes ~., data = train_qda)
qda_pred = predict(qda_stand, train_qda)
pred_val_qda = qda_pred$class
test_error_qda = mean(pred_val_qda != true_val)
cat("Test error rate of model is: ", test_error_qda)
```
**From normal qda process, we get an error rate of 0.509, which is relevantly high**
```{r}
# Apply PCA
pca_train <- prcomp(train_qda, rank = 6)$x
pca_train <- as.data.frame(pca_train)
pca_test = prcomp(test_qda, rank = 6)$x
pca_test <- as.data.frame(pca_test)
# QDA with PCA
qda_model_pca = qda(classes ~., data = pca_train, method = "mle")
train_pred_pca = predict(qda_model_pca, pca_train)$class
train_error = mean(train_pred_pca != train_qda$classes)
test_pred_pca = predict(qda_model_pca, pca_test)$class
test_error = mean(test_pred_pca != test_qda$classes)
cat("Train error rate: ", train_error, "\n")
cat("Test error rate: ", test_error)
```
**Quadratic Discriminant Analysis - Log transformed**

```{r}
# Read in the datasets
train_qda_log <- read.csv("train_log.csv")
train_qda_log <- subset(train_qda_log, select = -c(1))
test_qda_log <- read.csv("test_log.csv")
test_qda_log <- subset(test_qda_log, select = -c(1))
```

```{r}
# Train QDA
library("MASS")
true_val_log = test_qda_log$classes
qda_log = qda(classes ~., data = train_qda_log)
qda_pred_log = predict(qda_log, train_qda_log)
pred_val_qda_log = qda_pred_log$class
test_error_qda_log = mean(pred_val_qda_log != true_val_log)
cat("Test error rate of model is: ", test_error_qda_log)
```
**From normal qda process, we get an error rate of 0.509, which is relevantly high**
```{r}
# Apply PCA
pca_train_log <- prcomp(train_qda_log, rank = 10)$x
pca_train_log <- as.data.frame(pca_train_log)
pca_test_log = prcomp(test_qda_log, rank = 10)$x
pca_test_log <- as.data.frame(pca_test_log)
# QDA with PCA
qda_model_pca_log = qda(classes ~., data = pca_train_log, method = "mle")
train_pred_pca_log = predict(qda_model_pca_log, pca_train_log)$class
train_error_log = mean(train_pred_pca_log != train_qda_log$classes)
test_pred_pca_log = predict(qda_model_pca_log, pca_test_log)$class
test_error_log = mean(test_pred_pca_log != test_qda_log$classes)
cat("Train error rate: ", train_error_log, "\n")
cat("Test error rate: ", test_error_log)
```

**Quadratic Discriminant Analysis - Discretized**

```{r}
# Read in the datasets
train_qda_dis<- read.csv("train_dis.csv")
train_qda_dis <- subset(train_qda_dis, select = -c(1))
test_qda_dis <- read.csv("test_dis.csv")
test_qda_dis <- subset(test_qda_dis, select = -c(1))
```

```{r}

```

```{r}
# Train QDA
library("MASS")
true_val_dis = test_qda_dis$classes
qda_dis = qda(classes ~., data = train_qda_dis)
qda_pred_dis = predict(qda_dis, train_qda_dis)
pred_val_qda_dis = qda_pred_dis$class
test_error_qda_dis = mean(pred_val_qda_dis != true_val_dis)
cat("Test error rate of model is: ", test_error_qda_dis)
```
**From normal qda process, we get an error rate of 0.509, which is relevantly high**
```{r}
# Apply PCA
pca_train_log <- prcomp(train_qda_log, rank = 10)$x
pca_train_log <- as.data.frame(pca_train_log)
pca_test_log = prcomp(test_qda_log, rank = 10)$x
pca_test_log <- as.data.frame(pca_test_log)
# QDA with PCA
qda_model_pca_log = qda(classes ~., data = pca_train_log, method = "mle")
train_pred_pca_log = predict(qda_model_pca_log, pca_train_log)$class
train_error_log = mean(train_pred_pca_log != train_qda_log$classes)
test_pred_pca_log = predict(qda_model_pca_log, pca_test_log)$class
test_error_log = mean(test_pred_pca_log != test_qda_log$classes)
cat("Train error rate: ", train_error_log, "\n")
cat("Test error rate: ", test_error_log)
```


