---
title: "Final_LDA"
author: "AmyLiu"
date: "2023-03-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Use the standardized training data to train LDA. The LDA model with 57 variables results in a testing error of 0.1. We also try applying PCA before applying LDA, but a rank of 57 results in a testing error of about 0.2. After testing out different ranks, PCA results in a testing error of 0.18.**

```{r}
# read in data
library("MASS")
train = read.csv("train_stand.csv")
train <- subset(train, select = -c(X))
test = read.csv("test_stand.csv")
test <- subset(test, select = -c(X))
true = test$classes
```

```{r}
# train LDA
lda.model = lda(classes ~., data = train)
summary(lda.model)
```

```{r}
lda.predicted = predict(lda.model,test)$class
lda.error = mean(lda.predicted != true) # take mean of number of mismatching across all labels
cat("Test error rate of model is :", lda.error)
```


----
**LDA can possibly be improved by: (1) PCA (2) Oversample class 0 (3) regularization (4) cross validation to choose the right number of dimensions (5)remove variables that are highly correlated**


```{r}
# Try PCA
pca.train = prcomp(subset(train, select = -c(classes)), scale = TRUE, rank=12)$x
pca.train <- as.data.frame(pca.train)
classes = train$classes
pca.train$classes = classes # new df with 26 independent vars and 1 dependent var

lda.pca.model =lda(classes ~ ., data = pca.train, method = "mle")
summary(lda.pca.model)
```


```{r}
# prepare test data - PCA everything except classes
pca.test = prcomp(subset(test, select = -c(classes)), scale = TRUE, rank=12)$x
pca.test <- as.data.frame(pca.test)
pca.test$classes = true # new df with 26 independent vars and 1 dependent var

lda.pca.predicted = predict(lda.pca.model, pca.test)$class
lda.pca.error = mean(lda.pca.predicted != true) # take mean of number of mismatching across all labels
cat("Test error rate of model is :", lda.pca.error)
```

